Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-06-16 paper.qmd
Updated naming with convention

nldr_mnist1 <- tsne_mnist |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#984ea3') +
  interior_annotation("a", c(0.08, 0.9)) +
  theme(aspect.ratio = 1)

@@

nldr_mnist6 <- tsne_mnist2 |>
  ggplot(aes(x = tSNE1,
             y = tSNE2))+
  geom_point(alpha=0.1, size=1, colour='#999999') +
  interior_annotation("f", c(0.08, 0.9)) +
  theme(aspect.ratio = 1)

@@

nldr_mnist2 <- umap_mnist |>
  ggplot(aes(x = UMAP1,
             y = UMAP2)) +
  geom_point(alpha=0.1, size=1, colour='#a65628') +
  interior_annotation("b") +
  theme(aspect.ratio = 1)

@@

nldr_mnist3 <- phate_mnist |>
  ggplot(aes(x = PHATE1,
             y = PHATE2))+
  geom_point(alpha=0.1, size=1, colour='#377eb8') +
  interior_annotation("c", c(0.94, 0.92)) +
  theme(aspect.ratio = 1)

@@

nldr_mnist4 <- trimap_mnist |>
  ggplot(aes(x = TriMAP1,
             y = TriMAP2))+
  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
  interior_annotation("d", position = c(0.86, 0.92)) +
  theme(aspect.ratio = 1)

@@

nldr_mnist5 <- pacmap_mnist |>
  ggplot(aes(x = PaCMAP1,
             y = PaCMAP2))+
  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
  interior_annotation("e", c(0.08, 0.9)) +
  theme(aspect.ratio = 1)
```

@@

tsne_minst_scaled_temp <- tsne_minst_scaled |>
  rename(c("emb1" = "tSNE1",
           "emb2" = "tSNE2"))

tsne_minst_scaled <- tsne_minst_scaled |>
  mutate(cluster = if_else((tSNE1 <= 1) & (tSNE1 >= 0.78) & (tSNE2 <= 0.62) & (tSNE2 >= 0.39), "small_clust", "big_clust"))

trimesh_removed_mnist <- ggplot() + 
  geom_point(data = tsne_minst_scaled,
             aes(
               x = tSNE1,
               y = tSNE2,
               colour = cluster
             ),
             alpha=0.2)+
    geom_segment(data = tr_from_to_df_mnist, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000") +
  scale_color_manual(values=c("#999999",'#ff7f00')) +
  interior_annotation("a", c(0.08, 0.93), cex = 2) +
  theme(legend.position = "none")

tsne_plot_mnist_clust <- ggplot() + 
  geom_point(data = tsne_minst_scaled,
             aes(
               x = tSNE1,
               y = tSNE2,
               colour = cluster
             ),
             alpha=0.2)+
  scale_color_manual(values=c('#d9d9d9', '#ff7f00')) +
  interior_annotation("a", c(0.08, 0.93), cex = 2.5) +
  theme(legend.position = "none")

@@

highd_error_2d_mnist <- error_df |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = sqrt_row_wise_tot_error)) +
  geom_point(alpha=0.2) +
  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
  geom_rect(aes(xmin = 0.82, xmax = 0.92, ymin = 0.38, ymax = 0.58),
               fill = "transparent", colour = "#000000", linewidth = 1) +
  geom_rect(aes(xmin = 0.9, xmax = 1, ymin = 1.25, ymax = 1.45),
               fill = "transparent", colour = "#375cb1", linewidth = 1)  +
  geom_rect(aes(xmin = 0.35, xmax = 0.45, ymin = 0.55, ymax = 0.7),
               fill = "transparent", colour = "#008000", linewidth = 1)  +
  geom_rect(aes(xmin = 0.7, xmax = 0.8, ymin = 0.25, ymax = 0.4),
               fill = "transparent", colour = "#4c389a", linewidth = 1) +
  interior_annotation(label = "a", c(0.08, 0.93), cex = 2) +
  theme(legend.position = "none",
        aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0),
        panel.spacing = unit(0, "pt"),
        axis.ticks.length = unit(0, "pt"),
        strip.text = element_blank(),   # Optional, if you're using facets
        axis.title = element_blank()   # Optional
        )
```

>>>>

nldr_mnist1 <- tsne_mnist |>
  ggplot(aes(x = emb1,
             y = emb2))+
  geom_point(alpha=0.1, size=1, colour='#984ea3') +
  interior_annotation("a", c(0.08, 0.9)) +
  theme(aspect.ratio = 1)

@@

nldr_mnist6 <- tsne_mnist2 |>
  ggplot(aes(x = emb1,
             y = emb2))+
  geom_point(alpha=0.1, size=1, colour='#999999') +
  interior_annotation("f", c(0.08, 0.9)) +
  theme(aspect.ratio = 1)

@@

nldr_mnist2 <- umap_mnist |>
  ggplot(aes(x = emb1,
             y = emb2)) +
  geom_point(alpha=0.1, size=1, colour='#a65628') +
  interior_annotation("b") +
  theme(aspect.ratio = 1)

@@

nldr_mnist3 <- phate_mnist |>
  ggplot(aes(x = emb1,
             y = emb2))+
  geom_point(alpha=0.1, size=1, colour='#377eb8') +
  interior_annotation("c", c(0.94, 0.92)) +
  theme(aspect.ratio = 1)

@@

nldr_mnist4 <- trimap_mnist |>
  ggplot(aes(x = emb1,
             y = emb2))+
  geom_point(alpha=0.1, size=1, colour='#ff7f00') +
  interior_annotation("d", position = c(0.86, 0.92)) +
  theme(aspect.ratio = 1)

@@

nldr_mnist5 <- pacmap_mnist |>
  ggplot(aes(x = emb1,
             y = emb2))+
  geom_point(alpha=0.1, size=1, colour='#e41a1c') +
  interior_annotation("e", c(0.08, 0.9)) +
  theme(aspect.ratio = 1)
```

@@

tsne_minst_scaled <- tsne_minst_scaled |>
  mutate(cluster = if_else((emb1 <= 1) & (emb1 >= 0.78) & (emb2 <= 0.62) & (emb2 >= 0.39), "small_clust", "big_clust"))

trimesh_removed_mnist <- ggplot() + 
  geom_point(data = tsne_minst_scaled,
             aes(
               x = emb1,
               y = emb2,
               colour = cluster
             ),
             alpha=0.2)+
    geom_segment(data = tr_from_to_df_mnist, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000") +
  scale_color_manual(values=c("#999999",'#ff7f00')) +
  interior_annotation("a", c(0.08, 0.93), cex = 2) +
  theme(legend.position = "none")

tsne_plot_mnist_clust <- ggplot() + 
  geom_point(data = tsne_minst_scaled,
             aes(
               x = emb1,
               y = emb2,
               colour = cluster
             ),
             alpha=0.2)+
  scale_color_manual(values=c('#d9d9d9', '#ff7f00')) +
  interior_annotation("a", c(0.08, 0.93), cex = 2.5) +
  theme(legend.position = "none")

@@

highd_error_2d_mnist <- error_df |>
  ggplot(aes(x = emb1,
             y = emb2,
             colour = sqrt_row_wise_tot_error)) +
  geom_point(alpha=0.2) +
  scale_colour_continuous_sequential(palette = "YlOrRd", n_interp = 20) +
  geom_rect(aes(xmin = 0.82, xmax = 0.92, ymin = 0.38, ymax = 0.58),
               fill = "transparent", colour = "#000000", linewidth = 1) +
  geom_rect(aes(xmin = 0.9, xmax = 1, ymin = 1.25, ymax = 1.45),
               fill = "transparent", colour = "#375cb1", linewidth = 1)  +
  geom_rect(aes(xmin = 0.35, xmax = 0.45, ymin = 0.55, ymax = 0.7),
               fill = "transparent", colour = "#008000", linewidth = 1)  +
  geom_rect(aes(xmin = 0.7, xmax = 0.8, ymin = 0.25, ymax = 0.4),
               fill = "transparent", colour = "#4c389a", linewidth = 1) +
  interior_annotation(label = "a", c(0.08, 0.93), cex = 2) +
  theme(legend.position = "none",
        aspect.ratio = 1,
        plot.margin = margin(0, 0, 0, 0),
        panel.spacing = unit(0, "pt"),
        axis.ticks.length = unit(0, "pt"),
        strip.text = element_blank(),   # Optional, if you're using facets
        axis.title = element_blank()   # Optional
        )
```
